---
title: "Boar Analysis Markdown"
author: "Sophie Fox-Gmuer"
date: "2024-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Binomial data; TB in wild boar

# The data: Variables
1. Tb: a binomial variable, categorising individuals as infected (1) or uninfected (2)
 
2. Sex: a categorical variable, defining individuals as male (1) or female (2) 

3. Age: a categorical variable defining individuals into 1 of 4 age categories - juvenile, subadult, young adult, old adult (accurate age cannot be clearly defined in years)

4. Length: nose-rump length, measured in cm

# The aim: identify the risk factors for TB in this population
 e.g. is infection equally distributed among the population? or can we make predictions about likelihood
```{r , echo=FALSE}

dframe1<-read.csv(choose.files(),na="NA",stringsAsFactors =T) # 

dframe1$age <- as.factor(dframe1$age)   
dframe1$sex <- as.factor(dframe1$sex)
plot(dframe1$age, xlab= "age") 
plot(dframe1$sex, xlab ="sex")

```


## GLM model 1

You can also embed plots, for example:

```{r , echo=FALSE}
hist(dframe1$length) 
model1 <- glm(Tb ~ length + sex + length:sex,family = binomial (link = logit), data = dframe1)
plot(model1)

```


## After testing for model efficiency this is the final GLM model
```{r , echo=FALSE}
library(effects) 

par(mfrow=c(1,1)) 

plot(allEffects(model1))
modelfinal <- glm(Tb ~ length,
              family = binomial (link = logit),
              data = dframe1)
plot(modelfinal)


```

## Visualise the effects
```{r , echo=FALSE}

library(effects)
par(mfrow=c(1,1)) 
plot(allEffects(model1)) 


```


## the final model
```{r , echo=FALSE}

modelfinal <- glm(Tb ~ length,
              family = binomial (link = logit),
              data = dframe1)

par (mfrow = c (1,1)) 
plot (dframe1$Tb ~ dframe1$length, 	
      ylab = "Probability of having Tb",
      xlab = "Body length (cm)" ,
      col = "blue")
pdat <- expand.grid(length = seq(40,180, 1) ) 
pdat 

pred <- predict (modelfinal, newdata = pdat, type= "response", se.fit = TRUE) 

predframe <- data.frame (pdat, preds = pred$fit, se = pred$se.fit)
predframe

lines (predframe$preds ~ predframe$length, col="red", lwd = 2)

predframe$upperse <- (predframe$preds + predframe$se )
lines (predframe$upperse ~ predframe$length, lty = 2, col = "red", lwd = 1.5) 

predframe$lowerse <- (predframe$preds - predframe$se)
lines (predframe$lowerse ~ predframe$length, lty = 2, col = "red", lwd = 1.5) 



predframe$upperCI <- (predframe$preds + (1.96*predframe$se ))
lines (predframe$upperCI ~ predframe$length, lty = 2, col = "blue", lwd = 2) 

predframe$lowerCI <- (predframe$preds - (1.96*predframe$se))
lines (predframe$lowerCI ~ predframe$length, lty = 2, col = "blue", lwd = 2) 


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
